---
title: "P8105 HW1"
output: github_document
author: Xinyu Yang
---
This is the R Markdown for P8150 Homework 1   
<br>

## Problem 1
<br>
**Create data frame**
```{r}
#load the tidyverse package
library(tidyverse)

#Set seed for random sampling
set.seed(123)

#Create data frame p1_df
p1_df = tibble(
  ran_sample = rnorm(8),
  vec_char = c("This", "is", "an", "example", "of", "the", "char", "vector"),
  vec_logical = ran_sample > 0,
  vec_factor = factor(c("Yes", "No", "Maybe", "Maybe", "Yes", "No", "Yes", "No")) 
)

```
<br> 

**Compute mean for each variable in data frame**
```{r}
#Compute means
mean(pull(p1_df, ran_sample))
mean(pull(p1_df, vec_char))
mean(pull(p1_df, vec_logical))
mean(pull(p1_df, vec_factor))

```
As seen above, among the 4 variables, mean can be computed for the random sample and the logical vector, but cannot be computed for the character or factor vector, which returns 'NA' as result.  
<br>   

**Apply as.numeric and convert**
```{r, results = 'hide', message = FALSE, warning = FALSE}

as.numeric(pull(p1_df, vec_char))
as.numeric(pull(p1_df, vec_logical))
as.numeric(pull(p1_df, vec_factor))

```
When 'as.numeric' is applied, character vector returns 'NA' by coercion, while logical and factor vectors return numeric values. This indicates that logical and factor vectors can be coerced into numeric values but character vector can't. This helps explain why logical vector can be used to compute mean, but doesn't explain why factor vector can't.  


```{r}
#convert the logical vector to numeric, and multiply the random sample by the result
convert_1 = as.numeric(pull(p1_df, vec_logical))
convert_1 * pull(p1_df, ran_sample)

#convert the logical vector to a factor, and multiply the random sample by the result
convert_2 = as.factor(pull(p1_df, vec_logical))
convert_2 * pull(p1_df, ran_sample)

#convert the logical vector to a factor and then convert the result to numeric, and multiply the random sample by the result
convert_3 = as.factor(pull(p1_df, vec_logical))
convert_3 = as.numeric(convert_3)
convert_3 * pull(p1_df, ran_sample)

```
As seen above, when the logical vector is ultimately converted to numeric, it can be multiplied by the random sample; however, when it is only converted into a factor, it cannot be multiplied by the random sample.

---

## Problem 2
<br>
**Create data frame**
```{r}

set.seed(12345)

p2_df = tibble(
  x = rnorm(500),
  y = rnorm(500),
  z_logic = (x + y > 1),
  z_numeric = as.numeric(z_logic),
  z = as.factor(z_numeric)
)

```
* The dataset contains `r nrow(p2_df)` rows and `r ncol(p2_df)` columns.  
* For sample x, the mean is `r round(mean(pull(p2_df, x)), digit=2)`, the median is `r round(median(pull(p2_df, x)), digit=2)`, and the standard deviation is `r round(sd(pull(p2_df, x)), digit=2)`
* `r sum(pull(p2_df, z_numeric)) * 100 / nrow(p2_df)`% of sample x satisfy that x + y > 1  

<br>

**Scatterplots of Y vs X**  

_Scatterplot 1_: group by logical variable (see below)
```{r}
#Generate the first scatterplot, color points using the logical variable
ggplot(p2_df, aes(x = x, y = y, color = z_logic)) + geom_point()

#Export the scatterplot to project directory
ggsave("scatter_plot.pdf", height = 4, width = 6)

```

<br>
  _Scatterplot 2_: group by numeric variable (see below)
```{r}
#Generate the second scatterplot, color points using the numeric variable
ggplot(p2_df, aes(x = x, y = y, color = z_numeric)) + geom_point()

```

<br>
  _Scatterplot 3_: group by factor variable (see below)
```{r}
#Generate the third scatterplot, color points using the factor variable
ggplot(p2_df, aes(x = x, y = y, color = z)) + geom_point()

```
  

  
